\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{enumitem}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{booktabs}

\lstset{
  language=Verilog,
  basicstyle=\ttfamily\small,
  keywordstyle=\color{blue}\bfseries,
  commentstyle=\color{gray},
  breaklines=true,
  frame=single,
  numbers=left,
  numberstyle=\tiny\color{gray}
}

\title{\textbf{Lab 5 --- Coverage}}
\author{}
\date{}

\begin{document}
\maketitle

\section*{FIFO}

\subsection*{Task 1}

\subsubsection*{Question 1.1: Covergroup Bins}

The testbench defines one \texttt{covergroup} (\texttt{fifo\_cov}) and one \texttt{cover property} directive. In total there are \textbf{13 covergroup bins + 1 cover directive}.

\paragraph*{Coverpoint \texttt{full} (1 bin)}
\begin{enumerate}[label=\arabic*.]
  \item \texttt{went\_full} --- bin for \texttt{full == 1}
\end{enumerate}

\paragraph*{Coverpoint \texttt{empty} (1 bin)}
\begin{enumerate}[resume, label=\arabic*.]
  \item \texttt{went\_empty} --- bin for \texttt{empty == 1}
\end{enumerate}

\paragraph*{Coverpoint \texttt{dut.count} (3 bins)}
\begin{enumerate}[resume, label=\arabic*.]
  \item \texttt{empty\_bin} --- occupancy equals 0
  \item \texttt{mid\_bins} --- occupancy in range \texttt{[1 : DEPTH-1]}
  \item \texttt{full\_bin} --- occupancy equals \texttt{DEPTH}
\end{enumerate}

\paragraph*{Coverpoint \texttt{wr\_en} (2 auto-bins)}
\begin{enumerate}[resume, label=\arabic*.]
  \item \texttt{auto[0]} --- \texttt{wr\_en == 0}
  \item \texttt{auto[1]} --- \texttt{wr\_en == 1}
\end{enumerate}

\paragraph*{Coverpoint \texttt{rd\_en} (2 auto-bins)}
\begin{enumerate}[resume, label=\arabic*.]
  \item \texttt{auto[0]} --- \texttt{rd\_en == 0}
  \item \texttt{auto[1]} --- \texttt{rd\_en == 1}
\end{enumerate}

\paragraph*{Cross \texttt{wr\_en} $\times$ \texttt{rd\_en} (4 bins)}
\begin{enumerate}[resume, label=\arabic*.]
  \item \texttt{write\_only} --- \texttt{wr\_en=1, rd\_en=0}
  \item \texttt{read\_only} --- \texttt{wr\_en=0, rd\_en=1}
  \item \texttt{both\_off} --- \texttt{wr\_en=0, rd\_en=0}
  \item \texttt{illegal} --- \texttt{wr\_en=1, rd\_en=1}
\end{enumerate}

\paragraph*{Cover Directive (+1)}
\begin{enumerate}[resume, label={D1.}]
  \item \texttt{cover property (becomes\_empty\_after\_read)} --- covers the sequence where the FIFO transitions to empty after a read when \texttt{dut.count > 0}.
\end{enumerate}

\subsubsection*{Question 1.2: Bins Covered After Reset Only}

The following \textbf{5 bins} are covered just from reset:

\begin{enumerate}
  \item \texttt{went\_empty} --- after reset the FIFO is empty, so \texttt{empty == 1}
  \item \texttt{empty\_bin} --- \texttt{dut.count == 0} after reset
  \item \texttt{wr\_en auto[0]} --- \texttt{wr\_en} stays 0
  \item \texttt{rd\_en auto[0]} --- \texttt{rd\_en} stays 0
  \item \texttt{both\_off} --- cross bin where \texttt{wr\_en=0, rd\_en=0}
\end{enumerate}

\subsection*{Task 2}

\subsubsection*{Question 2.1: Additionally Covered Bins}

After enabling the provided stimulus, the following \textbf{5 additional bins} are now covered:

\begin{enumerate}
  \item \texttt{write\_only} --- cross bin where \texttt{wr\_en=1, rd\_en=0}
  \item \texttt{read\_only} --- cross bin where \texttt{wr\_en=0, rd\_en=1}
  \item \texttt{wr\_en auto[1]} --- \texttt{wr\_en == 1}
  \item \texttt{rd\_en auto[1]} --- \texttt{rd\_en == 1}
  \item \texttt{mid\_bins} --- occupancy in range \texttt{[1 : DEPTH-1]}
\end{enumerate}

\subsection*{Task 3}

\subsubsection*{Missing Bins After Task 2}

The following \textbf{3 bins + 1 cover directive} remained uncovered:

\begin{enumerate}
  \item \texttt{went\_full} --- \texttt{full == 1}
  \item \texttt{full\_bin} --- \texttt{dut.count == DEPTH}
  \item \texttt{illegal} --- cross bin where \texttt{wr\_en=1, rd\_en=1}
\end{enumerate}
\begin{enumerate}[label={D1.}]
  \item \texttt{becomes\_empty\_after\_read} --- FIFO transitions to empty after a read when \texttt{dut.count > 0}
\end{enumerate}

\subsubsection*{Additional Stimulus}

\paragraph*{Fully fill and fully empty the FIFO.}
Increased the number of consecutive writes to exceed \texttt{DEPTH} (16), driving the FIFO to full occupancy, then increased the number of consecutive reads to drain it completely back to empty.

\textbf{Hits bins:}
\begin{itemize}
  \item \texttt{went\_full} --- FIFO reaches full capacity
  \item \texttt{full\_bin} --- \texttt{dut.count == DEPTH}
  \item \texttt{becomes\_empty\_after\_read} (D1) --- FIFO drains to empty after a read while \texttt{dut.count > 0}
\end{itemize}

\paragraph*{Drive both \texttt{wr\_en} and \texttt{rd\_en} on the same cycle.}
Added stimulus that asserts \texttt{wr\_en=1} and \texttt{rd\_en=1} simultaneously.

\textbf{Hits bin:}
\begin{itemize}
  \item \texttt{illegal} --- cross bin where \texttt{wr\_en=1, rd\_en=1}
\end{itemize}

\section*{Cache}

\subsection*{Task 4}

\subsubsection*{Covergroup Bins}

The testbench defines one \texttt{covergroup} (\texttt{cache\_cov}) with \textbf{9 coverpoints/crosses} totalling \textbf{27 bins}, plus \textbf{7 cover property} directives.

\paragraph*{Coverpoint \texttt{write} (2 auto-bins)}
\begin{enumerate}[label=\arabic*.]
  \item \texttt{auto[0]} --- \texttt{write == 0}
  \item \texttt{auto[1]} --- \texttt{write == 1}
\end{enumerate}

\paragraph*{Coverpoint \texttt{read} (2 auto-bins)}
\begin{enumerate}[resume, label=\arabic*.]
  \item \texttt{auto[0]} --- \texttt{read == 0}
  \item \texttt{auto[1]} --- \texttt{read == 1}
\end{enumerate}

\paragraph*{Coverpoint \texttt{hit} (2 auto-bins)}
\begin{enumerate}[resume, label=\arabic*.]
  \item \texttt{auto[0]} --- \texttt{hit == 0}
  \item \texttt{auto[1]} --- \texttt{hit == 1}
\end{enumerate}

\paragraph*{Cross \texttt{write} $\times$ \texttt{read} (4 bins)}
\begin{enumerate}[resume, label=\arabic*.]
  \item \texttt{write\_only} --- \texttt{write=1, read=0}
  \item \texttt{read\_only} --- \texttt{write=0, read=1}
  \item \texttt{both\_off} --- \texttt{write=0, read=0}
  \item \texttt{read\_over\_write} --- \texttt{write=1, read=1}
\end{enumerate}

\paragraph*{Cross \texttt{write} $\times$ \texttt{hit} (4 bins)}
\begin{enumerate}[resume, label=\arabic*.]
  \item \texttt{write\_hit} --- \texttt{write=1, hit=1}
  \item \texttt{write\_miss} --- \texttt{write=1, hit=0}
  \item \texttt{auto[write=0][hit=1]} --- not writing, hit asserted (carried from prior op)
  \item \texttt{auto[write=0][hit=0]} --- not writing, no hit
\end{enumerate}

\paragraph*{Cross \texttt{read} $\times$ \texttt{hit} (4 bins)}
\begin{enumerate}[resume, label=\arabic*.]
  \item \texttt{read\_hit} --- \texttt{read=1, hit=1}
  \item \texttt{read\_miss} --- \texttt{read=1, hit=0}
  \item \texttt{auto[read=0][hit=1]} --- not reading, hit asserted (carried from prior op)
  \item \texttt{auto[read=0][hit=0]} --- not reading, no hit
\end{enumerate}

\paragraph*{Coverpoint \texttt{dut.tag} (3 bins)}
\begin{enumerate}[resume, label=\arabic*.]
  \item \texttt{lower\_bins} --- tag in range \texttt{[0 : TAG\_LO]}
  \item \texttt{mid\_bins} --- tag in range \texttt{[TAG\_LO+1 : TAG\_HI-1]}
  \item \texttt{high\_bins} --- tag in range \texttt{[TAG\_HI : TAG\_MAX]}
\end{enumerate}

\paragraph*{Coverpoint \texttt{dut.index} (3 bins)}
\begin{enumerate}[resume, label=\arabic*.]
  \item \texttt{lower\_bins} --- index in range \texttt{[0 : INDEX\_LO]}
  \item \texttt{mid\_bins} --- index in range \texttt{[INDEX\_LO+1 : INDEX\_HI-1]}
  \item \texttt{high\_bins} --- index in range \texttt{[INDEX\_HI : INDEX\_MAX]}
\end{enumerate}

\paragraph*{Coverpoint \texttt{dut.offset} (3 bins)}
\begin{enumerate}[resume, label=\arabic*.]
  \item \texttt{lower\_bins} --- offset in range \texttt{[0 : OFFSET\_LO]}
  \item \texttt{mid\_bins} --- offset in range \texttt{[OFFSET\_LO+1 : OFFSET\_HI-1]}
  \item \texttt{high\_bins} --- offset in range \texttt{[OFFSET\_HI : OFFSET\_MAX]}
\end{enumerate}

\paragraph*{Cover Directives (+7)}
\begin{enumerate}[label={D\arabic*.}]
  \item \texttt{p\_read\_replace} --- read to a valid line with tag mismatch (conflict miss / replacement)
  \item \texttt{p\_write\_replace} --- write to a valid line with tag mismatch (conflict miss / replacement)
  \item \texttt{p\_read\_cold\_miss} --- read to an invalid line (compulsory miss)
  \item \texttt{p\_write\_cold\_miss} --- write to an invalid line (compulsory miss)
  \item \texttt{p\_write\_then\_read\_hit} --- write followed by read hit to the same address within 5 cycles (data round-trip)
  \item \texttt{p\_read\_miss\_then\_hit} --- read miss followed by read hit to the same address within 5 cycles (cache fill path)
  \item \texttt{p\_hit\_deasserts} --- hit transitions $1 \rightarrow 0$ (verifies hit is not stuck high)
\end{enumerate}

\subsubsection*{Uncovered Items with Simple Random Stimulus (\texttt{cache2})}

The \texttt{cache2} testbench uses plain \texttt{\$urandom\_range(0,255)} for address and a strict 50/50 read-or-write split over 200 cycles. This leaves \textbf{1 covergroup bin} and \textbf{2 cover directives} uncovered:

\begin{enumerate}
  \item \texttt{read\_over\_write} (cross \texttt{write} $\times$ \texttt{read}) --- the stimulus always asserts exactly one of \texttt{read} or \texttt{write}. \texttt{read=1, write=1} is structurally impossible with this stimulus.
  \item \texttt{p\_write\_then\_read\_hit} --- requires a write followed by a read to the same address within 5 cycles. With 256 possible addresses, the chance of the same address repeating in that window is too low to guarantee coverage over 200 cycles.
  \item \texttt{p\_read\_miss\_then\_hit} --- same reason; requires same-address repetition that pure random stimulus does not reliably produce.
\end{enumerate}

\subsubsection*{Changes Made to Achieve Full Coverage (\texttt{cache})}

Two changes were made to the stimulus and one to the cover properties relative to \texttt{cache2}:

\paragraph*{1. Weighted constrained-random operation class.}
Replaced the 50/50 \texttt{\$urandom\_range} with a SystemVerilog transaction class using a \texttt{dist} constraint, enabling all four operation combinations to be generated:

\begin{center}
\begin{tabular}{lrl}
  \toprule
  Operation & Weight & Signals \\
  \midrule
  Read only    & 45\% & \texttt{read=1, write=0} \\
  Write only   & 45\% & \texttt{read=0, write=1} \\
  Simultaneous &  5\% & \texttt{read=1, write=1} \\
  Idle         &  5\% & \texttt{read=0, write=0} \\
  \bottomrule
\end{tabular}
\end{center}

\textbf{Covers:} \texttt{read\_over\_write} --- the 5\% simultaneous weight guarantees \texttt{read=1, write=1} is generated within 200 cycles.

\paragraph*{2. Directed sequences for same-address access.}
Two directed loops were added after the random phase:

\begin{itemize}
  \item \textit{Write then read-hit:} writes to a random address, then reads the same address on the next cycle. Repeated 20 times. Covers \texttt{p\_write\_then\_read\_hit}.
  \item \textit{Read-miss then read-hit:} reads a random address (miss), then reads the same address on the next cycle (guaranteed hit). Repeated 20 times. Covers \texttt{p\_read\_miss\_then\_hit}.
\end{itemize}

\paragraph*{3. Relaxed temporal constraint on sequential properties.}
Both \texttt{p\_write\_then\_read\_hit} and \texttt{p\_read\_miss\_then\_hit} were updated from \texttt{\#\#1} to \texttt{\#\#[1:5]}, allowing the matching access to occur within a 5-cycle window rather than requiring it on the immediately following cycle.

\end{document}
